Active Intent Disambiguation for \\ Shared Autonomy Robots

### 2019-07-28 LatexStuff_Revision_2/Reviews (copy)/information_on_response_to_reviewers/paper_review_response_template/reviewresponse_example.tex

\subsection{Task Success and Task Completion Times}
As per the reviewers' suggestion we analyzed the task completion times and we are reporting the results. As our algorithm was not time optimized and also due to the lack of transparency regarding \textit{why} the autonomy is choosing the disambiguation modes, a statistically significant difference was not observed between the manual and disambiguation. However, an overall trend of a reduction in variance was observed in the the disambiguating trials indicating that the task performance was in general more consistent. We also report the overall task success (94.53$\%$).

\rcomment{
 Comparison with memory-based intent inference and Recursive belief method is not convincing....
}

\textbf{Response}

The reviewer's comment inspired us to look even more closely into the inner workings of the field-theoretic approach and conduct simulation-based quantitative evaluation of the different intent inference approaches. 
A new section has been added to discuss the advantages of field-theoretic approach in the context of assistive robotics.

@@

A key requirement of an intent inference algorithm in the context of assistive setting is that it should be able reason about goal probabilities in the \textbf{absence} of useful information. In our system we rely solely on user control commands as our evidence variable for inference. Subjects may not issue velocity control commands for large chunks of time due to fatigue, motor impairment, mode switches, cognitive load etc.

We show that in the absence of control commands our field-theoretic approach converges according to the \textit{principle of maximum entropy} to a uniform distribution whereas recursive Bayesian approaches as implemented in [23] converges to the stationary distribution of the underlying goal transition matrix which can potentially create unwanted biases in the inference.

We have replaced the qualitative comparison with a simulation-based quantitative comparison of memory-based, recursive Bayesian and the field-theoretic approach. The inference accuracy for field-theoretic approach was comparable to the recursive Bayesian updating but significantly better than the memory-based approach indicating that no degradation in inference accuracy when the control commands are available.

\section{Reviewer 2}

We are glad that you were happy with the submitted version. In this revised version we have further improved different sections of the paper according to the suggestion of Reviewer 1 and we believe that this has significantly strengthened our paper.

### 2019-07-28 LatexStuff_Revision_2/bare_jrnl.tex
Final -1

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
Assistive shared-control robots have the potential to transform the lives of millions of people afflicted with severe motor impairments. The usefulness of shared-control robots typically relies on the underlying autonomy's ability to infer the user's needs and intentions, and the ability to do so \textit{unambiguously} is often a limiting factor for providing appropriate assistance confidently and accurately. The contributions of this paper are three-fold. First, we propose a \textit{goal disambiguation} algorithm that enhances the intent inference and assistive capabilities of a shared-control assistive robotic arm. Second, we introduce a field-theoretic approach to \textit{intent inference} inspired by dynamic field theory that works in tandem with the disambiguation scheme. Third, we present a pilot study with eight subjects to evaluate the efficacy of the disambiguation algorithm. Our results suggest that (a) the disambiguation system is of greater utility for more limited control interfaces and more complex tasks, and (b) subjects demonstrated a wide range of disambiguation request behaviors with the common thread of concentrating requests early in the execution.
%(c) there are no differences in the onset of the robot assistance between different mode switching paradigms across tasks or across interfaces. 
\end{abstract}

@@

In this section, we propose a novel intent inference scheme inspired by \textit{dynamic field theory} in which the time evolution of the probability distribution $\boldsymbol{p}(t)$ is specified as a dynamical system with constraints. Section~\ref{ssec:dft} provides a primer on the basic principles and features of \textit{dynamic field theory} and its application in the fields of neuroscience and cognitive robotics. \added{Section~\ref{ssec:dft_ii} describes our formulation that makes use of dynamic field theory for the purposes of intent inference (Section~\ref{ssec:dft_adv}) presents the characteristics of the formulation and its applicability to assistive robotics and a simulation-based quantitative comparison of different goal inference methods. }
\subsection{Dynamic Field Theory}\label{ssec:dft}

@@

\subsection{\added{Field-Theoretic Intent Inference For Assistive Robotics}}\label{ssec:dft_adv}
\begin{figure}[b]
	\centering
	\includegraphics[width = 1.0\hsize]{./figures/inference_comparison_new.eps}
%	\vspace{-1cm}
	\caption{\added{Goal inference comparison: (Left to Right). Goal probabilities for Memory-Based Prediction (Left), Recursive Bayesian Belief Update (Middle) and DFT-based inference scheme (Right) and reaching behavior in a scene with two goals. The black boxes in the plots indicated time range during which the control velocity was zeroed. The probabilities do not change much for memory-based approach as the cost function employed is purely distance-based (left). The convergence is towards the stationary and uniform distribution for recursive (middle) and field-theoretic approach (right) respectively.}}
	\label{fig:dft_inference}
\end{figure}
\added{In our work, autonomy's inference of user intent solely relies on user control commands. In the domain of assistive robotics, it is quite often the case that the user input is highly discontinuous (due to fatigue, motor impairments, stoppage for mode switches \textit{et cetera}) due to which it becomes important to reason about belief over goals in the \textit{absence} of useful information. }

\added{In the absence of testable information (that is, no control commands issued by the user and not task-level global priors), in accordance to the \textit{maximum entropy principle}, the belief should converge to a uniform distribution. In the absence of $\boldsymbol{u}_h$ Equation~\ref{eq:dft_ii} converges to a uniform distribution mimicking natural forgetting behavior resulting in a more meaningful inference. The rate at which the distribution evolves to a uniform distribution is controlled by the timescale parameter $\tau_1$.}

\added{On the other hand, standard discrete-time recursive belief update equation as implemented in~\cite{jain2018recursive} is}
\begin{equation*}
p(g^t |\boldsymbol{u}_h^t) = \eta p(\boldsymbol{u}_h^t | g^t)\sum_{g^{t-1} \in \mathcal{G}}^{}p(g^t | g^{t-1})p(g^{t-1} |\boldsymbol{u}_h^{t-1})\;\;\;\;\;.
\end{equation*}
\added{where $\eta$ is the normalization factor, $\boldsymbol{p}(\boldsymbol{u}_h | g)$ is likelihood function, $p(g^t | g^{t-1})$ is the goal transition probability. 
In recursive belief update when $\boldsymbol{u}_h = 0$ with the likelihood function to be uniform, it can be shown that the posterior distribution over goals converges to the stationary distribution of the goal transition matrix. The stationary distribution is not necessarily uniform and therefore is counter-intuitive and can introduce unwanted biases in the inference. 
}

\added{
Knowledge of task-level semantics can provide informative global priors that can further improve the accuracy of the inference mechanism. Our field-theoretic approach can encode a task-level global prior in the `rest state' term. For example, in a multi-stage task such as pouring the contents of a cup into a bowl, the initial goal distribution will be biased towards the cup as it is more likely the target that the user would try to reach for. The relative values of $\tau_1$ and $\tau_2$ controls the convergence behavior (whether to uniform distribution or rest state distribution respectively) in the absence of $\boldsymbol{u}_h$.}

@@

\added{In order to evaluate the performance of the field-theoretic inference approach a point robot simulation-based quantitative comparison to 1) memory-based prediction~\cite{dragan2013policy} and 2) recursive belief updating~\cite{jain2018recursive} was implemented in $\mathbb{R}^3$. The human is modeled as issuing a control command that noisily optimizes a straight-line path towards the intended goal. Signal dropout was simulated by randomly zeroing out control commands. Additionally, $\boldsymbol{u}_h$ was set to be zero for a randomly chosen section of each trial in order to compare the convergence behavior of different approaches. The number of goals varied between three and five and goal transitions were randomly sampled every five to eight time steps. 500 trials were simulated. Inference accuracy is computed as the fraction of total trial time (excluding when $\boldsymbol{u}_h = 0$) for which an algorithm correctly inferred the ground truth. Results for DFT-based inference were comparable to recursive belief updating ($87.46\%$ and $87.43\%$ respectively) and outperformed memory-based prediction significantly ($59.15\%$). Figure~\ref{fig:dft_inference} shows an illustrative example of goal inference using the various methods. One can see that when there is no control command issued unlike the recursive belief approach, the field-theoretic approach converges to the maximum entropy uniform distribution.}

@@

\added{Additionally, at the end of each testing phase, subjective data was gathered via a brief questionnaire. Users were given the following statements regarding the usefulness and capability of the assistance system to rate according to their agreement.
}
\begin{itemize}
	\item \textbf{Q1} - \added{Control modes chosen by the system made task execution easier.}
	\item \textbf{Q2} - \added{I liked operating the robots in the control modes chosen by the system.}
	\item \textbf{Q3} - \added{The robot and I worked together to accomplish the task.}
	\item \textbf{Q4} - \added{Which interface was the hardest to operate? 1-Joystick, 2-Head array.}
	\item \textbf{Q5} - \added{For which interface was the assistance paradigm the most useful 1-Joystick, 2-Head array.}
	\item \textbf{Q6} - \added{Which one of the schemes do you prefer the most? 1-Manual, 2-Disambiguation}
	\item \textbf{Q7} - \added{Which one of the schemes is the most user-friendly? 1-Manual, 2-Disambiguation}
\end{itemize}
\added{A 7-point Likert scale was used to rate \textbf{Q1-3} and for \textbf{Q4-Q7} subjects were required to pick one of the two given options. }

@@

\noindent{\textbf{Task Completion Times:}} 
\added{No statistical difference was observed in task completion times between the \textit{Manual} and \textit{Disambiguation} paradigm. However, the variance in the task completion times was lower in \textit{Disambiguation} paradigm indicating that the task performance was more consistent when robot control was performed in the disambiguating control modes chosen by the autonomy. Furthermore, the overall task success was $94.53\%$ (242 out of 256 trials). 10 out of the 14 unsuccessful trials happened during the multi-step task.}
\begin{figure}[h!]
	\centering
	\includegraphics[width = 1\hsize, height=0.35\vsize, ,left]{./figures/task_completion.eps}
	\caption{Task completion times between \textit{Disambiguation} and \textit{Manual} Paradigms. \textit{Left:} Grouped by control interfaces. \textit{Right:} Grouped by tasks.}
	\label{fig:task_completion}
\end{figure}

@@

This highlights the need for greater transparency in the human-robot interaction, so that the human has a clear picture of \textit{how} and \textit{why} the robot chooses to help the user in certain ways. A lack of understanding of how they might help the robot to help them might have resulted in the under-utilization of the disambiguation feature. In order to provide \textit{intent-expressive} control commands to the robot, very likely knowledge of the assistance mechanism is critical.  
%\added{Important to note is that this work was not about having the system select a control mode that the \textit{human} wants to use to reach a target (which might be motivated by a wide variety of factors such as line of sight or ease of use). In spite of this, subject responses to the user survey indicated that majority of the subjects liked operating the robot in the disambiguating modes and felt that task execution became easier as a result.} 
Therefore, the need for extensive and thorough training becomes apparent.
The training can be made more effective in a few different ways. First, online feedback of the robot's intent prediction at all times during training can likely help the subject gain a better understanding of the relationship between the characteristics of their control actions (sparsity, aggressiveness, persistence) and the robot's assistive behavior. Second, the subjects could be explicitly informed of the task relevant features (directedness, proximity \textit{et cetera}) that the robot relies on for determining the amount of assistance. Knowledge of these features might motivate the users to leverage the disambiguating mode.

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
Assistive shared-control robots have the potential to transform the lives of millions of people afflicted with severe motor impairments. The usefulness of shared-control robots typically relies on the underlying autonomy's ability to infer the user's needs and intentions, and the ability to do so \textit{unambiguously} is often a limiting factor for providing appropriate assistance confidently and accurately. The contributions of this paper are three-fold. First, we propose a \textit{goal disambiguation} algorithm that enhances the intent inference and assistive capabilities of a shared-control assistive robotic arm. Second, we introduce a field-theoretic approach to \textit{intent inference} inspired by dynamic field theory that works in tandem with the disambiguation scheme. Third, we present a pilot study with eight subjects to evaluate the efficacy of the disambiguation algorithm. Our results suggest that (a) the disambiguation system \added{helped to significantly reduce task effort as measured by number of button presses} and is of greater utility for more limited control interfaces and more complex tasks, and (b) subjects demonstrated a wide range of disambiguation request behaviors with the common thread of concentrating requests early in the execution.
%(c) there are no differences in the onset of the robot assistance between different mode switching paradigms across tasks or across interfaces. 
\end{abstract}

@@

In this section, we propose a novel intent inference scheme inspired by \textit{dynamic field theory} in which the time evolution of the probability distribution $\boldsymbol{p}(t)$ is specified as a dynamical system with constraints. Section~\ref{ssec:dft} provides a primer on the basic principles and features of \textit{dynamic field theory} and its application in the fields of neuroscience and cognitive robotics. \added{Section~\ref{ssec:dft_ii} describes our formulation that makes use of dynamic field theory for the purposes of intent inference. Section~\ref{ssec:dft_adv} presents the characteristics of the formulation, its suitability to assistive robotics and a simulation-based quantitative comparison of different goal inference methods. }
\subsection{Dynamic Field Theory}\label{ssec:dft}

@@

\subsection{\added{Field-Theoretic Intent Inference For Assistive Robotics}}\label{ssec:dft_adv}
\begin{figure}[b]
	\centering
	\includegraphics[width = 1.0\hsize]{./figures/inference_comparison_new.eps}
%	\vspace{-1cm}
	\caption{\added{Goal inference comparison: (Left to Right). Goal probabilities for Memory-Based Prediction (Left), Recursive Bayesian Belief Update (Middle) and field-theoretic inference scheme (Right) and reaching behavior in a scene with two goals. The black rectangular boxes in the plots indicates time range during which the control velocity was zeroed. The probabilities do not change much for memory-based approach as the cost function employed is purely distance-based (left). The convergence is towards the stationary and uniform distribution for recursive (middle) and field-theoretic approach (right) respectively.}}
	\label{fig:dft_inference}
\end{figure}
\added{In our work, autonomy's inference of user intent solely relies on user control commands. In the domain of assistive robotics, it is quite often the case that the user input is highly discontinuous (due to fatigue, motor impairments, stoppage for mode switches \textit{et cetera}). Therefore, it is important to reason about belief over goals in the \textit{absence} of useful information. }

\added{According to the \textit{principle of maximum entropy}, in the absence of testable information (no control commands issued by the user and uniform global prior), the belief should converge to a uniform distribution. In the absence of $\boldsymbol{u}_h$, using Equation~\ref{eq:dft_ii}, $\boldsymbol{p}(t)$ converges to a uniform distribution mimicking natural forgetting behavior and results in a more meaningful inference. The rate at which the distribution decays to a uniform distribution is controlled by $\tau_1$.}

\added{On the other hand, standard discrete-time recursive belief update equation as implemented in~\cite{jain2018recursive} is}
\begin{equation*}
p(g^t |\boldsymbol{u}_h^t) = \eta p(\boldsymbol{u}_h^t | g^t)\sum_{g^{t-1} \in \mathcal{G}}^{}p(g^t | g^{t-1})p(g^{t-1} |\boldsymbol{u}_h^{t-1})\;\;\;\;\;.
\end{equation*}
\added{where $\eta$ is the normalization factor, $\boldsymbol{p}(\boldsymbol{u}_h | g)$ is likelihood function, $p(g^t | g^{t-1})$ is the goal transition probability. 
In recursive belief update when $\boldsymbol{u}_h = 0$, with the likelihood function to be uniform, it can be shown that the posterior distribution over goals converges to the stationary distribution of the goal transition matrix. The stationary distribution is not necessarily uniform and can introduce unwanted biases in the inference. 
}

\added{
Knowledge of task-level semantics can provide informative global priors that can further improve the accuracy of the inference mechanism. Our field-theoretic approach can encode a task-level global prior in the `rest state' term. For example, in a multi-stage task such as pouring the contents of a cup into a bowl, the initial goal distribution will be biased towards the cup as it is more likely  that the user would try to reach for the cup initially. The relative values of $\tau_1$ and $\tau_2$ control the convergence behavior (whether to uniform distribution or rest state distribution respectively) in the absence of $\boldsymbol{u}_h$.}

@@

\added{In order to evaluate the performance of the field-theoretic inference approach a point robot simulation-based quantitative comparison to 1) memory-based prediction~\cite{dragan2013policy} and 2) recursive belief updating~\cite{jain2018recursive} was implemented in $\mathbb{R}^3$. The human is modeled as issuing a control command that noisily optimizes a straight-line path towards the intended goal. Signal dropout was simulated by randomly zeroing out control commands. Additionally, $\boldsymbol{u}_h$ was set to be zero for a randomly chosen section of each trial in order to compare the convergence behavior of different approaches. The number of goals varied between three and five and goal transitions were randomly sampled every five to eight time steps. 500 trials were simulated. Inference accuracy was computed as the fraction of total trial time (excluding when $\boldsymbol{u}_h = 0$) for which an algorithm correctly inferred the ground truth. Results for field-theoretic inference were comparable to recursive belief updating ($87.46\%$ and $87.43\%$ respectively) and outperformed memory-based prediction significantly ($59.15\%$). Figure~\ref{fig:dft_inference} shows an illustrative example of goal inference using the various methods. One can see that when there is no control command issued unlike the recursive belief approach, the field-theoretic approach converged to the maximum entropy uniform distribution.}

@@

\added{Additionally, at the end of each testing phase, subjective data was gathered via a brief questionnaire. Users were given the following statements regarding the usefulness and capability of the assistance system to rate according to their agreement.
}
\begin{itemize}
	\item \textbf{Q1} - \added{Control modes chosen by the system made task execution easier.}
	\item \textbf{Q2} - \added{I liked operating the robots in the control modes chosen by the system.}
	\item \textbf{Q3} - \added{The robot and I worked together to accomplish the task.}
	\item \textbf{Q4} - \added{Which interface was the hardest to operate? 1-Joystick, 2-Head array.}
	\item \textbf{Q5} - \added{For which interface was the assistance paradigm the most useful 1-Joystick, 2-Head array.}
	\item \textbf{Q6} - \added{Which one of the schemes do you prefer the most? 1-Manual, 2-Disambiguation}
	\item \textbf{Q7} - \added{Which one of the schemes is the most user-friendly? 1-Manual, 2-Disambiguation}
\end{itemize}
\added{A 7-point Likert scale was used to rate \textbf{Q1-3} and for \textbf{Q4-Q7} subjects were required to choose one of the two given options. }

@@

\noindent{\textbf{Task Completion Times:}} 
\added{No statistical difference was observed in task completion times between the \textit{Manual} and \textit{Disambiguation} paradigm. However, the variance in the task completion times was lower in \textit{Disambiguation} paradigm (except for Joystick/SingleStep) indicating that the task performance was generally more consistent when robot control was performed in the disambiguating control modes chosen by the autonomy. Furthermore, the overall task success was $94.53\%$ (242 out of 256 trials). 10 out of the 14 unsuccessful trials happened during the multi-step task.}
\begin{figure}[h!]
	\centering
	\includegraphics[width = 1\hsize, height=0.35\vsize, ,left]{./figures/task_completion.eps}
	\caption{Task completion times between \textit{Disambiguation} and \textit{Manual} Paradigms. \textit{Left:} Grouped by control interfaces. \textit{Right:} Grouped by tasks.}
	\label{fig:task_completion}
\end{figure}

@@

This highlights the need for greater transparency in the human-robot interaction, so that the human has a clear picture of \textit{how} and \textit{why} the robot chooses to help the user in certain ways. A lack of understanding of how they might help the robot to help them might have resulted in the under-utilization of the disambiguation feature. In order to provide \textit{intent-expressive} control commands to the robot, very likely, knowledge of the assistance mechanism is critical.  
%\added{Important to note is that this work was not about having the system select a control mode that the \textit{human} wants to use to reach a target (which might be motivated by a wide variety of factors such as line of sight or ease of use). In spite of this, subject responses to the user survey indicated that majority of the subjects liked operating the robot in the disambiguating modes and felt that task execution became easier as a result.} 
Therefore, the need for extensive and thorough training becomes apparent.
The training can be made more effective in a few different ways. First, online feedback of the robot's intent prediction at all times during training can likely help the subject gain a better understanding of the relationship between the characteristics of their control actions (sparsity, aggressiveness, persistence) and the robot's assistive behavior. Second, the subjects could be explicitly informed of the task relevant features (directedness, proximity \textit{et cetera}) that the robot relies on for determining the amount of assistance. Knowledge of these features might motivate the users to leverage the disambiguating mode.
